Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit

### 2024-05-29 optim.qmd
print the smoothness and squintability values in the pdf

```{r setup, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(patchwork)
```

@@

[We generate random points that is beyond 1.5 projection distance and interpolate. Then we fit a kernel or use nonlinear least squares.]

@@

[Nicolas's pdf]

Second, we consider the product of the largest absolute magnitude of rate of change of $f$ and the corresponding projection angle as a second measure of squintability. Since $f$ is decreasing, the rate of change of $f$ is negative and thus \[|\underset{x}{\min} f('x)|\] gives the absolute magnitude of the most negative rate of change.

[Nicolas's pdf]

@@

[example plots]

@@

>>>>

```{r setup, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggh4x)
```

@@

\[We generate random points that is beyond 1.5 projection distance and
interpolate. Then we fit a kernel or use nonlinear least squares.\]

@@

\[Nicolas's pdf\]

Second, we consider the product of the largest absolute magnitude of
rate of change of $f$ and the corresponding projection angle as a second
measure of squintability. Since $f$ is decreasing, the rate of change of
$f$ is negative and thus $|\underset{x}{\min} f('x)|$ gives the absolute
magnitude of the most negative rate of change.

\[Nicolas's pdf\]

@@

\[example plots\]

@@

construct a relationship between jellyfish success and jellyfish
parameters and the optimisation properties defined in @sec-theory.

This can imform the choice of jellyfish parameters for a given
optimisation problem.

In addition to the pipe-finding problem, we also consider the problem of
finding the sine wave structure in the 6D space and six indexes
(`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`) are
considered. Combined with the jellyfish optimisers (`n_jellies` =
20/50/100, `max_tries` = 50/100), we obtain additional 27 setups for
this investigation.

```{r}
load(here::here("data/smoothness.rda"))
smoothness |> select(index, n, variance:nugget) |> knitr::kable(digits = 3)
load(here::here("data/squintability.rda"))
squintability |> knitr::kable(digits = 3)
```

```{r}
#| fig.width = 8,
#| fig.height = 5,
#| fig.cap = "sdkflskdfjl"
load(here::here("data/sq_basis_dist_idx.rda"))
sq_basis_dist_idx |>
  ggplot(aes(x = dist, y = idx, group = index)) +
  geom_line() +
  facet_nested_wrap(~index + n, nrow = 3, labeller = label_both, scales = "free_y") + 
  xlab("projection distance") + ylab("index value")

### 2024-05-29 optim.tex
print the smoothness and squintability values in the pdf

@@

\begin{tcolorbox}[enhanced jigsaw, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, breakable, titlerule=0mm, opacityback=0, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomrule=.15mm, left=2mm, coltitle=black]

@@

From the literature, it is commonly understood that a large squint angle
implies that the function is easy to optimize, because we do not need to
be very close to the perfect view to see the structure. A small squint
angle means that the derivative of the index function can still be very
large values near the optimal point and will rapidly change as we get
even closer to the optimum. As such we can observe the second order
gradient, which is the rate of change of gradient, over the space we are
searching. For some index functions, the second order gradient is not
well-defined, we can approximate the second order gradient vector in
similar fashion as the above section.

@@

>>>>

\begin{verbatim}
Warning: package 'ggh4x' was built under R version 4.3.2
\end{verbatim}

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, toprule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, arc=.35mm, breakable, leftrule=.75mm, opacitybacktitle=0.6, bottomtitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, bottomrule=.15mm, left=2mm, opacityback=0]

@@

From the literature, it is commonly understood that a large squint angle
implies that the objective function value is close to optimal even when
we are not very close to the perfect view to see the structure. A small
squint angle means that index function value improves substantially only
when we are very close to the perfect view. As such, low squintability
implies rapid improvement in the index value when near the perfect view.

In this study, we propose two metrics to capture the notion of
squintability.

{[}We generate random points that is beyond 1.5 projection distance and
interpolate. Then we fit a kernel or use nonlinear least squares.{]}

First, parametric model.

{[}Nicolas's pdf{]}

Second, we consider the product of the largest absolute magnitude of
rate of change of \(f\) and the corresponding projection angle as a
second measure of squintability. Since \(f\) is decreasing, the rate of
change of \(f\) is negative and thus \(|\underset{x}{\min} f('x)|\)
gives the absolute magnitude of the most negative rate of change.

{[}Nicolas's pdf{]}

@@

construct a relationship between jellyfish success and jellyfish
parameters and the optimisation properties defined in
Section~\ref{sec-theory}.

This can imform the choice of jellyfish parameters for a given
optimisation problem.

In addition to the pipe-finding problem, we also consider the problem of
finding the sine wave structure in the 6D space and six indexes
(\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}) are considered. Combined with the
jellyfish optimisers (\texttt{n\_jellies} = 20/50/100,
\texttt{max\_tries} = 50/100), we obtain additional 27 setups for this
investigation.

\begin{longtable}[]{@{}lrrrrr@{}}
\toprule\noalign{}
index & n & variance & range & smoothness & nugget \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 \\
loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 \\
MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 \\
TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 \\
stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 \\
splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 \\
MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 \\
TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 \\
holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 \\
holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 \\
holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 \\
holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 \\
\end{longtable}

\begin{longtable}[]{@{}lrrrrr@{}}
\toprule\noalign{}
index & n & theta1 & theta2 & theta3 & theta4 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
holes & 6 & 1.001 & 0.860 & 3.368 & 0.823 \\
holes & 8 & 1.001 & 0.869 & 3.264 & 0.811 \\
holes & 10 & 1.000 & 0.885 & 3.151 & 0.806 \\
holes & 12 & 1.000 & 0.878 & 3.345 & 0.806 \\
MIC & 6 & 0.894 & 0.571 & 1.623 & -0.024 \\
MIC & 8 & 0.932 & 0.328 & 1.314 & -0.030 \\
TIC & 6 & 0.951 & 0.536 & 1.719 & -0.025 \\
TIC & 8 & 0.945 & 0.564 & 1.723 & -0.027 \\
dcor2d\_2 & 6 & 0.954 & 1.039 & 2.742 & -0.019 \\
loess2d & 6 & 1.016 & 1.039 & 2.648 & 0.080 \\
splines2d & 6 & 1.014 & 1.051 & 2.730 & -0.009 \\
stringy & 6 & 1.011 & 0.011 & 254.734 & 0.727 \\
\end{longtable}

{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-10-1.pdf}

\caption{sdkflskdfjl}
